---
title: "Homework 1"
author: "Conall Gordon 17390566"
format: pdf
editor: visual
execute: 
  eval: true
  echo: false
  message: false
  warning: false
---

Below I will introduce the data set which is available in base R, named '*trees*'. I will then investigate the data by performing a simple numerical summary and a basic plot. This file has been shared publicly on GitHub.

## Introduction to "Trees" Data

The '*trees*' data consists of the measurements of **31 felled black cherry trees**. The measurements recorded include the **diameter** (labelled '**girth**'), **height** and **volume** of the timber. The girth of each tree is measured at 4 ft 6 in above the ground.

The following units were used for each measurement:

-   **Height:** feet

-   **Girth:** inches

-   **Volume:** cubic feet

## Data Exploration

Below I perform a basic exploration of the data. The first rows can be seen below:

```{r}
#| label: Load and print data head

data("trees") #Load data

head(trees) #Print first rows of dataset

```

\newpage

A numerical summary of the data:

```{r}
#| label: Numerical summary

summary(trees) #Print simple numerical summary
```

We can see the numerical summary above. The following observations can be made:

-   **Girth:**

    -   The values **span** from 8.30 to 20.60 inches.

    -   The **mean** value is 13.25 inches.

-   **Height:**

    -   The values **span** from 63 to 87 feet.

    -   The **mean** value is 76 feet.

-   **Volume:**

    -   The values **span** from 10.20 to 77.00 cubic feet.

    -   The **mean** value is 30.17 cubic feet.

Note that the volume measurements span a wide range relative to the other measurements, likely due to the dimensionality of the volume measurement.

\newpage

## Scatter Plot

Below, I plot the **volume** of the felled trees against their **girth**. This plot allows for the visualization of the relationship between the volume and girth of the trees.

```{r}
#| label: Scatter plot of data
#| fig.cap: "Scatter plot of Tree Volume against Tree Girth"


#Load the ggplot2 package
library(ggplot2)

#Scatter plot using ggplot2
ggplot(trees, aes(x = Girth, y = Volume)) +
  geom_point(color = "red", shape = 19) + #Solid circle in red
  labs(
    title = "Scatter plot of Volume vs. Tree Girth", #Title
    x = "Girth (inches)",  #x-label
    y = "Volume (cubic feet)"  #y-label
  ) +
  scale_x_continuous(limits = c(0, max(trees$Girth))) + #Begin x-axis at 0
  scale_y_continuous(limits = c(0, max(trees$Volume))) + #Begin y-axis at 0
  theme_minimal() #Set theme


```

We can see above that there is a **clear positive relationship** between the **volume** of the tree timber and the **girth**. We see that as the girth increases, the volume can also be seen to increase. This relationship would be **expected**. This relationship could be explored further, for example the presence of a linear or quadratic relationship could be investigated.
